\begin{vita}

The back matter environment is the vita.  The vita environment
provides information on the author.  The content should focus on the
author's professional career.

Captain Ronald B. Cole graduated from Our Lady of Good Counsel High School in Wheaton, Maryland. He entered undergraduate studies at the University of Maryland in College Park, Maryland where he graduated with a Bachelor of Science degree in Marketing in May 1996. He was commissioned through the Detachment 330 AFROTC at the University of Maryland where he was recognized as a Distinguished Graduate and nominated for a Regular Commission.

His first assignment was at Columbus AFB as a student in Undergraduate Pilot Training in February 1997. In May 1998, he was assigned to the 99th Contracting Squadron, Nellis AFB, Nevada where he served as an operational acquisition contracting officer. While stationed at Nellis, he deployed overseas in July 2000 to spend three months in Riyadh, Saudi Arabia as the Joint Task Force-Southwest Asia contingency contracting officer. In May 2002, he entered the Graduate School of Engineering and Management, Air Force Institute of Technology. Upon graduation, he will be assigned to the Pentagon.

a huge note:

Note: It is prohibited to include any personal information in the following categories about U.S. citizens, DOD Employees and military personnel: social security account numbers; home addresses; dates of birth; telephone numbers other than duty officers which are appropriately made available to the general public; names, locations and any other identifying information about family members of DOD employees and military personnel. See AFI 33-332 for more information.

\end{vita}



 



